---
title: "Curve Fitting"
output: html_notebook
---

$$
 P(t) = p_{min} + \frac{p_{max}-p_{min}}{1 + e^{-r.(t-s)}}
$$

```{r}
richard <- function(t, p_max, p_min, r, s){
  return(p_min + (p_max - p_min)  / (1 + exp(-r * (t - s))))
}

```

\$\$ \begin{align*}
P(t) &= p_{min} + \frac{p_{max} - p_{min}}
                      {1 + e^{-r \cdot (t - s)}} \\

P'(t) &= -\frac{(p_{min} - p_{max}) \cdot r \cdot e^{r \cdot (t - s)}}
              {(e^{r \cdot (t - s)} + 1)^2} \\

P''(t) &= -\frac{(p_{min} - p_{max}) \cdot r^2 \cdot (e^{r \cdot (t - s)} - 1) \cdot e^{r \cdot (t - s)} }
               {(e^{r \cdot (t - s)} + 1)^3} \\

P'''(t) &= -\frac{(p_{min} - p_{max}) \cdot r^3 \cdot e^{r \cdot (t - s)} \cdot (e^{2r \cdot (t - s)} - 4e^{r \cdot (t - s}) + 1}
                {(e^{r \cdot (t - s)} + 1)^4} \\
\end{align*} \$\$

## Replace $r$ in $P(t)$ by $r_{max}$ which is the extremum of $P'(t)$ *i.e.* the value of $P'(t)$ where $P''(t) = 0$

$$
-\frac{(p_{min} - p_{max}) . r^2 . (e^{r . (t - s)} - 1) . e^{r . (t - s)} } {(e^{r . (t - s)} + 1)^3} \stackrel{?}{=} 0
$$

### Case 1 : $p_{min} - p_{max} = 0$

$$
\begin{align*}
p_{min} - p_{max} &= 0 \\
p_{min} &= p_{max}
\end{align*}
$$

Result in no growth

### Case 2 : $r^2 = 0$

$$
\begin{align*}
r^2 &= 0 \\
r &= 0
\end{align*}
$$

Result in no growth

### Case 3 : $e^{r(t-s)} = 0$

$$
\begin{align*}
e^{r \cdot (t - s)} &= 0
\end{align*}
$$

Impossible

### Case 4 : $e^{r(t-s)} -1 = 0$

$$
\begin{align*}
e^{r \cdot (t - s)} - 1 &= 0 \\
e^{r \cdot (t - s)} &= 1 \\
r \cdot (t - s) &= 0 \\
\end{align*}
$$

$$
r = 0
$$

Result in no growth

$$
t = s
$$

Ok

## Determine whether the extremum is a minima, a maxima or 0

Replace $t$ by $s$ in $P'''(t)$

\$\$ \begin{align*}
P'''(t) &= -\frac{(p_{min} - p_{max}) \cdot r^3 \cdot e^{r \cdot (t - s)} \cdot (e^{2r \cdot (t - s)} - 4e^{r \cdot (t - s)} + 1)}
                {(e^{r \cdot (t - s)} + 1)^4} \\

P'''(s) &= -\frac{(p_{min} - p_{max}) \cdot r^3 \cdot e^{r \cdot (s - s)} \cdot (e^{2r \cdot (s - s)} - 4e^{r \cdot (s - s)} + 1)}
                {(e^{r \cdot (s - s)} + 1)^4} \\

P'''(s) &= -\frac{(p_{min} - p_{max}) \cdot r^3 \cdot 1 \cdot (1 - 4 + 1)}
                {(1 + 1)^4} \\

P'''(s) &= -\frac{-2r^3 \cdot (p_{min} - p_{max}) }
                {16} \\
\end{align*} \$\$

### Case 1 : $r > 0$

$$
P'''(S) < 0 \\
P'(S) \rightarrow maximum
$$

### Case 2 : $r < 0$

$$
P'''(S) > 0 \\
P'(S) \rightarrow minimum
$$

### Case 3 : $r = 0$

$$
P'''(S) < 0 \\
P'(S) \rightarrow 0
$$

Let's find the value of the extremum *i.e.* $P'(s)$

\$\$ \begin{align*}

P'(t) &= -\frac{(p_{min} - p_{max}) \cdot r \cdot e^{r \cdot (t - s)}}
              {(e^{r \cdot (t - s)} + 1)^2} \\

P'(s) &= -\frac{(p_{min} - p_{max}) \cdot r \cdot e^{r \cdot (s - s)}}
              {(e^{r \cdot (s - s)} + 1)^2} \\

\frac{-4P'(s)}{(p_{min} - p_{max})} &= r \\

\frac{-4r_{max}}{(p_{min} - p_{max})} &= r \\


\end{align*} \$\$

Let's then replace $r$ in $P(t)$ by its value

$$
\begin{align*}
 P(t) &= p_{min} + \frac{p_{max}-p_{min}}{1 + e^{-r.(t-s)}} \\
 P(t) &= p_{min} + \frac{p_{max}-p_{min}}{1 + e^{4r_{max}.(t-s)/p_{min}- p_{max}}}
\end{align*}
$$

```{r}
richard2 <- function(t, p_max, p_min, r_max, s){
  return(p_min + (p_max - p_min)  / (1 + exp(4 * r_max * (t - s)/(p_min - p_max))))
}
```

We can now use this new equation to fit our date using a genetic algorithm

```{r}
fit_richard2 <- function(y, t) {
  
  max_y <- max(y)
  min_y <- min(y)
  r_boundary <- plyr::round_any(max_y-min_y,
                      10^(floor(log10(max_y-min_y))),
                      f = ceiling)
  r_boundary <- if_else(is.na(r_boundary), 0, r_boundary)
  s_boundary <- max(t)
  sign <- sign(max_y)

  
  fit <- GA::ga(type = "real-valued",
         fitness = \(p) -sum((y - richard2(t = t,  p[1], p[2], p[3], p[4]))^2),
         lower   = c(
           p_max = max_y - 0.01*max_y * sign,
           p_min = min_y - 0.01*max_y * sign,
           r_max = -r_boundary,
           s     = 0),
         upper = c(
           p_max = max_y + 0.01*max_y  * sign,
           p_min = min_y + 0.01*max_y  * sign,
           r_max = r_boundary,
           s     = s_boundary),
         names = c(
           "p_max",
           "p_min",
           "r_max",
           "s"),
         popSize = 100,
         monitor = FALSE)
  return(fit@solution[1,])
}
```

## Load some sample data

```{r}
test_data_gfp <- readxl::read_excel(
  path = "/home/thomas/Téléchargements/21-07-03 GFP Gain 150/21-07-03 GFP Gain 150/21-07-03-GFP gain 150.xlsx",
  sheet = "GFP 150",
  na = "OVRFLW")

test_data_od <- readxl::read_excel(
  path = "/home/thomas/Téléchargements/21-07-03 GFP Gain 150/21-07-03 GFP Gain 150/21-07-03-GFP gain 150.xlsx",
  sheet = "OD600",
  na = "OVRFLW")
```

```{r}
bg_gfp <- 6000
bg_od <- 0.045
library(glue)
library(ggtext)
library(tidyverse)
library(fuzzyjoin)

test_data_gfp |> 
  select(Time, matches("^[A-Za-z]{1}\\d{1,2}")) |> 
  mutate(Time = (Time - first(Time)) |> as.numeric()/3600) |> 
  difference_inner_join(
    test_data_od |> 
    select(Time, matches("^[A-Za-z]{1}\\d{1,2}")) |> 
    mutate(Time = (Time - first(Time)) |> as.numeric()/3600),
    by = c("Time" = "Time"),
    max_dist = 0.01) |> 
  pivot_longer(cols = c(matches("^[A-Za-z]{1}\\d{1,2}\\.x"), matches("^[A-Za-z]{1}\\d{1,2}\\.y")),
               names_to = c("well_bdello", "well_coli"),
               names_sep = ".") |> 
  pivot_longer(cols = matches("^[A-Za-z]{1}\\d{1,2}\\.y"),
               names_to = "well_coli")
 
test_data_gfp |> 
  select(Time, matches("^[A-Za-z]{1}\\d{1,2}")) |> 
  mutate(Time = (Time - first(Time)) |> as.numeric()/3600) |> 
  pivot_longer(cols = matches("^[A-Za-z]{1}\\d{1,2}"),
               names_to = "well")  |> 
  mutate(value = value - {{bg_gfp}}) |> 
  drop_na(value) ->  cleaned_gfp

test_data_od |> 
  select(Time, matches("^[A-Za-z]{1}\\d{1,2}")) |> 
  mutate(Time = (Time - first(Time)) |> as.numeric()/3600) |> 
  pivot_longer(cols = matches("^[A-Za-z]{1}\\d{1,2}"),
               names_to = "well")  |> 
    mutate(value = value - {{bg_od}}) |> 
  drop_na(value) ->  cleaned_od


cleaned_gfp|> 
  group_by(well) |> 
  nest(data = c(Time, value)) |> 
  mutate(fit = purrr::map(data, ~fit_richard2(.$value, .$Time))) |> 
  unnest_wider(fit) |> 
  select(-data) -> parameters_gfp


cleaned_od|> 
  group_by(well) |> 
  nest(data = c(Time, value)) |> 
  mutate(fit = purrr::map(data, ~fit_richard2(.$value, .$Time))) |> 
  unnest_wider(fit) |> 
  select(-data) -> parameters_od

parameters_gfp |> 
  inner_join(parameters_od,
             by = c("well"= "well"),
             suffix = c("_bb", "_ec")) |> 
  inner_join(cleaned_od) |> 
  group_by(well) |> 
  mutate(
    r_max_bb = r_max_bb/(p_max_bb - p_min_bb),
    r_max_ec = r_max_ec/(p_max_ec - p_min_ec),
    across(c(p_max_bb,p_min_bb), .fns = ~(. - p_min_bb)/(p_max_bb - p_min_bb)),
    across(c(p_max_ec,p_min_ec), .fns = ~(. - p_min_ec)/(p_max_ec - p_min_ec)))|> 
  rowwise() |> 
  mutate(fit_bb = richard2(Time, p_max_bb, p_min_bb, r_max_bb, s_bb),
         fit_ec = richard2(Time, p_max_ec, p_min_ec, r_max_ec, s_ec),
         fit_sum = fit_ec - fit_bb)  |> 
  ggplot(aes(x = Time)) +
  geom_line(aes(y=fit_bb), color = "red") +
  geom_line(aes(y=fit_ec), color = "blue") +
  geom_line(aes(y=fit_bb), color = "red") +
  geom_line(aes(y=fit_ec), color = "blue") +
#  geom_line(aes(y=fit_sum), color = "green")+
  facet_wrap(~well)+
  theme_minimal() -> plt


  
cleaned_od |> arrange(well)
parameters |> 
  mutate(annot = glue("rate<sub>max</sub> {sprintf('%0.2f', r_max)}")) -> annot

cleaned_gfp |>
  inner_join(parameters) |> 
  arrange(well) |> 
  rowwise() |> 
  mutate(fit = richard2(Time, p_max, p_min, r_max, s),
#         max_r_t = ((log(abs(-(-p_min+(3*p_max)) / (p_min+p_max))) * (p_min-p_max))/(4*r_max))+s,
#         max_r_y = richard2(max_r_t, p_max, p_min, r_max, s),
#         test = max_r_y - ((p_max - p_min)/2 + p_min),
         annot = glue("rate<sub>max</sub> {sprintf('%0.2f', r_max)}")) |> 
  ggplot(aes(x = Time, y = value, group = well), color = "black") +
  geom_line() +
  geom_line(aes(x = Time, y = fit,  group = well), color = "red") +
 # geom_vline(aes(xintercept = max_r_t)) +
  geom_hline(aes(yintercept = p_max), color = "green") +
  geom_hline(aes(yintercept = p_min), color = "green") +
 # geom_richtext(aes(label = annot), x = 5, y = 4500, hjust = 1, vjust = 1) +
  facet_wrap(~well) +
  theme_minimal()
  

```

```{r}
svg("~/curves_normal.svg", width = 16, height = 10)
plt
dev.off()

```

$$
 t=\frac{\ln \left(\frac{A-Y}{Y-B}\right)\left(B-A\right)}{4r}+s \\
t=\frac{\ln \left(-\frac{-B+3A}{B+A}\right)\left(B-A\right)}{4r}+s
$$

```{r}
library(deSolve)

LV <- function(alpha, beta,sigma,gamma, N1i, N2i, r1, r2, times) {
  
  LV_equations <- function(time, variables, parameters) {
    with(as.list(c(variables, parameters)), {
      dN1 <- alpha*preys - (beta*preys*predators)
      dN2 <- sigma*preys*predators - (gamma*predators)  
      return(list(c(dN1, dN2)))
    })
  }
  
  parameters_values <- c(alpha=alpha,
                         beta=beta,
                         sigma=sigma,
                         gamma=gamma)
  
  initial_values <- c(preys=N1i, predators=N2i)
  
  out <- ode(initial_values, times, LV_equations, parameters_values)
  
  as.data.frame(out)
}

out <- LV(alpha=0,
          beta=0.25,
          sigma=0.205, gamma=0.3/24 ,N1i=1, N2i=0.1,times=seq(0,24,1))

plot(x=out[,1], y=out[,2], type = "l")
lines(x=out[,1], y=out[,3], col="red")


```
